---
output:
  pdf_document: default
  html_document: default
---
```{r}
library(readr)
library(tidyverse)
library(scales)
library(RColorBrewer)
library(ggthemes)
library(lubridate)
library(ggrepel)
library(reshape)
library(gridExtra)
library(maps)
library(stringr)
library(ggcorrplot)
library(viridis)
df <- read_csv("countries of the world.csv")
View(df)

```
```{r}
dim(df)
summary(df)
```
This dataset contains 227 rows and 20 cloumns.
110

```{r}
sum(is.na(df))
```
There are 110 null values present in the data set.
```{r}
sapply(df, function(x) sum(is.na(x)))
```

```{r}
colnames(df)
df <- df[,c("Country","Population","Area","PopDensity","GDP","Literacy","Phones","Birthrate","Deathrate","Agriculture","Industry","Service")]
names(df) <- c("Region","Population","Area","PopDensity","GDP","Phones","Literacy","Birthrate","Deathrate","Agriculture","Industry","Service")
```


```{r}
df$Region <- gsub(" ", "", df$Region)
df$PopDensity <- as.numeric(gsub(",", ".", df$PopDensity))
df$Phones <- as.numeric(gsub(",", ".", df$Phones))
df$Literacy <- as.numeric(gsub(",", ".", df$Literacy))
df$Birthrate	<- as.numeric(gsub(",", ".", df$Birthrate))
df$Deathrate	<- as.numeric(gsub(",", ".", df$Deathrate))
df$Agriculture <- as.numeric(gsub(",", ".", df$Agriculture))
df$Industry <- as.numeric(gsub(",", ".", df$Industry))
df$Service <- as.numeric(gsub(",", ".", df$Service))
```


```{r}
library(dplyr)
plot(df$GDP)
plot(df$Literacy)
plot(df$Phones)
plot(df$Birthrate)
plot(df$Deathrate)
plot(df$Agriculture)
plot(df$Service)
plot(df$Industry)
```
Replacing the null/missing values with their respective mean.
```{r}
df$GDP[is.na(df$GDP)] <- mean(df$GDP, na.rm = TRUE)
df$Literacy[is.na(df$Literacy)] <- mean(df$Literacy, na.rm = TRUE)
df$Phones[is.na(df$Phones)] <- mean(df$Phones, na.rm = TRUE)
df$Birthrate[is.na(df$Birthrate)] <- mean(df$Birthrate, na.rm = TRUE)
df$Deathrate[is.na(df$Deathrate)] <- mean(df$Deathrate, na.rm = TRUE)
df$Agriculture[is.na(df$Agriculture)] <- mean(df$Agriculture, na.rm = TRUE)
df$Service[is.na(df$Service)] <- mean(df$Service, na.rm = TRUE)
df$Industry[is.na(df$Industry)] <- mean(df$Industry, na.rm = TRUE)
sum(is.na(df))
view(df)
```
Successfully replaced all NA's and missing values present in the dataset with their respective mean.

```{r}
options(warn = -1)
options(scipen = 10000)
options(repr.plot.width = 13.8, repr.plot.height = 9.2)
library(ggcorrplot)
annotate <- ggplot2::annotate
core <- cor(df[,c(2:ncol(df))], method = "spearman", use = "complete.obs")
options(repr.plot.width = 13, repr.plot.height = 11.18)
ggcorrplot(core, outline.col = "gray30", type = "upper",
           legend.title = "Strength of \ncorrelation", colors = c("#6D9EC1", "white", "#E46726"))+
  labs(y = "", x = "", title = "Spearman's correlation matrix for continuous variables")
```

Positive correlations are displayed in  orange and negative correlations in blue color.
Color intensity and the size of the circle are proportional to the correlation coefficients.
(Area,Population),(Literacy,GDP),(Service,Literacy) are strongly correlated.
(Birthrate,Phones),(Birthrate,GDP),(Agriculture,GDP),(Agriculture,Phones) are negatively correlated.
