---
title: "EDA ON COUNTRY VACCINATIONS DATASET"
author: "PES1UG20CS622,    Dept.    of    CSE    -    ADITYA SUNDAR RAJ"
subtitle: PES    University
output:
  pdf_document: default
  html_document:
    df_print: paged
urlcolor: blue
editor_options: null
markdown: null
wrap: 72
---

```{r}
library(readr)
country_vaccinations <- read_csv("~/Desktop/DA_PROJECT/country_vaccinations.csv")
country_vaccinations <- country_vaccinations[,c("country", "total_vaccinations", "date", "people_vaccinated", "daily_vaccinations_raw","people_vaccinated_per_hundred", "daily_vaccinations_per_million", "vaccines")]
```

```{r}
dim(country_vaccinations)
sum(is.na(country_vaccinations))
summary(is.na(country_vaccinations))
sapply(country_vaccinations, function(x) sum(is.na(x)))
var1 <- unique(country_vaccinations[,c("country","date")])
dim(var1)
```

The data-set we are working on here has 86512 ROWS and 8 COLUMNS.

It has a very sizable number of missing values, here 184790 observations across the data-set.

Data inconsistency prevails as long as missing values are not treated properly.

Duplicates are also looked into and resolved due to the combined uniqueness of two attributes in this particular data-set



```{r}
library(lubridate)
country_vaccinations$date <- as.Date(country_vaccinations$date)
country_vaccinations$date <- as.Date(country_vaccinations$date)
country_vaccinations$total_vaccinations[is.na(country_vaccinations$total_vaccinations)==T] <- 0 
country_vaccinations$people_vaccinated[is.na(country_vaccinations$people_vaccinated)==T] <- 0 
country_vaccinations$daily_vaccinations_raw[is.na(country_vaccinations$daily_vaccinations_raw)==T] <- 0 
country_vaccinations$people_vaccinated_per_hundred[is.na(country_vaccinations$people_vaccinated_per_hundred)==T] <- 0 
country_vaccinations$daily_vaccinations_per_million[is.na(country_vaccinations$daily_vaccinations_per_million)==T] <- 0 
head <- country_vaccinations[sample(1:nrow(country_vaccinations),5), ]
head[order(head$date),]
country_vaccinations$month <- month(country_vaccinations$date)
country_vaccinations$weekday <- weekdays(country_vaccinations$date)
country_vaccinations$percent_people <- country_vaccinations$people_vaccinated_per_hundred/100
numcol_country_vaccinations <- country_vaccinations[,c('total_vaccinations','people_vaccinated','daily_vaccinations_raw','people_vaccinated_per_hundred','daily_vaccinations_per_million','month','percent_people')]
```


Missing values have been filled with zeroes as no other metric is suitable. 

This is done to ensure completeness and help us with our further observations.




```{r}
library(ggpubr)
library(corrplot)
M = cor(numcol_country_vaccinations)
corrplot(M, method = 'color')
```

The correlation plot can be observed to say there is no negative correlation between any of the attributes.

percent_people and people_vaccinated_per_hundred is very strongly correlated. 

Most attributes that depend on people or attributes that directly contribute to another attribute (eg: people_vaccinated and total_vaccinations) show high correlation.


COMMENTED CODE:

```{r}
#library(fpp2)
 #autoplot(ts(numcol_country_vaccinations$total_vaccinations))
 #autoplot(ts(numcol_country_vaccinations$people_vaccinated))
 #autoplot(ts(numcol_country_vaccinations$daily_vaccinations_raw))
 #autoplot(ts(numcol_country_vaccinations$people_vaccinated_per_hundred))
 #autoplot(ts(numcol_country_vaccinations$daily_vaccinations_per_million))
 #autoplot(ts(numcol_country_vaccinations$month))

 #autoplot(ts(numcol_country_vaccinations$percent_people))

 #tsoutliers(numcol_country_vaccinations$total_vaccinations)

 #tsoutliers(numcol_country_vaccinations$people_vaccinated)

#tsoutliers(numcol_country_vaccinations$daily_vaccinations_raw)

 #tsoutliers(numcol_country_vaccinations$people_vaccinated_per_hundred)

 #tsoutliers(numcol_country_vaccinations$daily_vaccinations_per_million)

 #tsoutliers(numcol_country_vaccinations$month)

 #tsoutliers(numcol_country_vaccinations$percent_people)


 #autoplot(tsclean(ts((numcol_country_vaccinations$total_vaccinations))), series="clean", color='red', lwd=0.9) +autolayer(ts(numcol_country_vaccinations), series="original", color='gray', lwd=1) +geom_point(data =tsoutliers(ts(numcol_country_vaccinations$total_vaccinations)) %>% as.data.frame(), aes(x=index, y=replacements), col='blue') +labs(x = "Time", y = "Total Vaccinations")


 #autoplot(tsclean(ts((numcol_country_vaccinations$people_vaccinated))), series="clean", color='red', lwd=0.9) +autolayer(ts(numcol_country_vaccinations), series="original", color='gray', lwd=1) +geom_point(data =tsoutliers(ts(numcol_country_vaccinations$people_vaccinated)) %>% as.data.frame(), aes(x=index, y=replacements), col='blue') +labs(x = "Time", y = "People Vaccinated")


 #autoplot(tsclean(ts((numcol_country_vaccinations$daily_vaccinations_raw))), series="clean", color='red', lwd=0.9) +autolayer(ts(numcol_country_vaccinations), series="original", color='gray', lwd=1) +geom_point(data = tsoutliers(ts(numcol_country_vaccinations$daily_vaccinations_raw)) %>% as.data.frame(),aes(x=index, y=replacements), col='blue') +labs(x = "Time", y = "Daily Vaccinations Raw")


 #autoplot(tsclean(ts((numcol_country_vaccinations$daily_vaccinations_raw))), series="clean", color='red', lwd=0.9) +autolayer(ts(numcol_country_vaccinations), series="original", color='gray', lwd=1) +geom_point(data = tsoutliers(ts(numcol_country_vaccinations$daily_vaccinations_raw)) %>% as.data.frame(), aes(x=index, y=replacements), col='blue') +labs(x = "Time", y = "Daily Vaccinations Raw")


 #autoplot(tsclean(ts((numcol_country_vaccinations$people_vaccinated_per_hundred))), series="clean", color='red', lwd=0.9) +autolayer(ts(numcol_country_vaccinations), series="original", color='gray', lwd=1) +geom_point(data = tsoutliers(ts(numcol_country_vaccinations$people_vaccinated_per_hundred)) %>% as.data.frame(), aes(x=index, y=replacements), col='blue') +labs(x = "Time", y = "People Vaccinated Per Hundred")


 #autoplot(tsclean(ts((numcol_country_vaccinations$daily_vaccinations_per_million))), series="clean", color='red', lwd=0.9) +autolayer(ts(numcol_country_vaccinations), series="original", color='gray', lwd=1) +geom_point(data = tsoutliers(ts(numcol_country_vaccinations$daily_vaccinations_per_million)) %>% as.data.frame(),aes(x=index, y=replacements), col='blue') +labs(x = "Time", y = "Daily Vaccinations Per Million")

 #autoplot(tsclean(ts((numcol_country_vaccinations$month))), series="clean", color='red', lwd=0.9) +autolayer(ts(numcol_country_vaccinations), series="original", color='gray', lwd=1) +geom_point(data = tsoutliers(ts(numcol_country_vaccinations$month)) %>% as.data.frame(),aes(x=index, y=replacements), col='blue') +labs(x = "Time", y = "Month")

#autoplot(tsclean(ts((numcol_country_vaccinations$percent_people))), series="clean", color='red', lwd=0.9) +autolayer(ts(numcol_country_vaccinations), series="original", color='gray', lwd=1) +geom_point(data = tsoutliers(ts(numcol_country_vaccinations$percent_people)) %>% as.data.frame(), aes(x=index, y=replacements), col='blue') +labs(x = "Time", y = "Percent People")

```


  A block of code has been commented above which identifies and caps the outliers that fall outside a certain range of values. 

CONCLUSION:

 Outliers were identified by transforming into time series data but could not be replaced by a suitable metric since this

 data-set comprises of real time data which is necessary for our study.

 Hence we will not be addressing them as outliers thus making the outlier count equal to 0.


```{r}
numcol_country_vaccinations.pca <- prcomp(numcol_country_vaccinations[,c(1:7)],
                   center = TRUE,
                   scale. = TRUE)

summary(numcol_country_vaccinations.pca)
```


 Proportion of variance for all 7 numeric principal components is low and PCA would not be the best option.

 Other transformations also do not seem fit due to the nature of this data-set.

```{r}
str(numcol_country_vaccinations.pca)
plot.numcol_country_vaccinations.pca <- plot(numcol_country_vaccinations.pca, type="l")
plot.numcol_country_vaccinations.pca
```

In the screeplot above, the ‘arm-bend’ represents a decrease in cumulative contribution. 

The above plot shows the bend at the third principal component.

```{r}
library(fpp2)
autoplot(ts(numcol_country_vaccinations$total_vaccinations))
autoplot(ts(numcol_country_vaccinations$people_vaccinated))
autoplot(ts(numcol_country_vaccinations$daily_vaccinations_raw))
autoplot(ts(numcol_country_vaccinations$people_vaccinated_per_hundred))
autoplot(ts(numcol_country_vaccinations$daily_vaccinations_per_million))
autoplot(ts(numcol_country_vaccinations$percent_people))
```


The plots above suggest that there are outliers but we will not be treating them for our analysis. 

This variation in values is what makes the base of our study.









